---
title: "Incidence Analysis Plots"
author: "Anie Thomas"
date: "2026-07-26"
output:
  html_document: default
  pdf_document: default
---

# Libraries
```{r}
#Load relevant libraries
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)
library(data.table)
library(sf)
library(stringr)
library(gtsummary)
library(gt)
library(patchwork)
library(writexl)
```

# Demographics table
```{r}
#Build demographics table
#Load up data
load("R:/Anie/cleaned_data_v2_JIA.RData")

#Building table
#Create separate data sets with valid age ranges
JIA_data <- data %>%
  filter(JIA == 1, JIA_age >= 0, JIA_age <= 16) %>% transmute(Disease = "JIA", AgeAtDiagnosis = JIA_age, gender, region)

JDM_data <- data %>%
  filter(JDM == 1, JDM_age >= 0, JDM_age <= 18) %>%
  transmute(Disease = "JDM", AgeAtDiagnosis = JDM_age, gender, region)

jSLE_data <- data %>%
  filter(jSLE == 1, jSLE_age >= 0, jSLE_age <= 18) %>%
  transmute(Disease = "jSLE", AgeAtDiagnosis = jSLE_age, gender, region)

#Combine datasets
demographic_data <- bind_rows(JIA_data, JDM_data, jSLE_data)

#Summary of data after filtering
summary(demographic_data)

demographic_data <- demographic_data %>%
  mutate(gender = recode(gender, "F" = "Female","M" = "Male"))

#Build demographics table
demographics_table <- demographic_data %>%
  tbl_summary(by = Disease, statistic = list(AgeAtDiagnosis ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"),label = list(AgeAtDiagnosis ~ "Age at diagnosis (years)", gender ~ "Sex", region ~ "Region"), missing = "no") %>% bold_labels() %>%
  modify_caption("**Demographic characteristics of incidence cases of JIA, JDM and jSLE**")

#View table
demographics_table

#Edit view of demographics table
demographics_table %>% as_gt() %>% tab_options(table.font.size = px(12))
```

#Population summary
```{r}
#Summary of total populations from all regions
table(data$region)
```
# Annual incidence by sex Line Graphs
## JDM
```{r}
#Load relevant .xlsx file
JDM_sex_graph <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual sex incidence rates.xlsx")
JDM_sex_graph$Group <- factor(JDM_sex_graph$Group, levels = c("F", "M")) 

#Plot JDM annual sex incidence
#2023 was excluded as there were no cases but large CI.
#Will have to make a table with 95% CI and add to supplementary/appendix
JDM_plot <- ggplot(
  JDM_sex_graph %>% filter(Year != 2023), aes(x = Year, y = IncidenceRate, colour = Group, group = Group)) +
  geom_line(linewidth = 1) + geom_point(size = 1) +  
  scale_colour_viridis_d(labels = c("F" = "Female", "M" = "Male")) +
  scale_x_continuous(breaks = seq(1995, 2022, by = 5)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years",
    colour = "Sex", title = "Annual incidence for JDM by sex (1995-2022)" ) +
  theme_classic(base_size = 12) +theme(axis.text = element_text(colour = "black"),
    legend.position = "right", legend.text = element_text(size = 11))

#View plot
JDM_plot

#Save plot to folder
#ggsave(JDM_plot, file = "R:/Anie/Incidence Plots/JDM/JDM_sex_incidence.png", width = 7, height = 5, dpi = 600)
```
## JIA
```{R}
#Load relevant xlsx data
JIA_sex_graph <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA Incidence rate - overall,sex,region.xlsx",
                            sheet = 2)
JIA_sex_graph$Group <- factor(JIA_sex_graph$Group, levels = c("F", "M")) 

#Plot JIA 
JIA_plot <- ggplot(JIA_sex_graph,
  aes(x = Year, y = IncidenceRate,colour = Group, group = Group)) +
  geom_line(linewidth = 1) + geom_point(size = 1.2) +
  scale_colour_viridis_d(labels = c("F" = "Female", "M" = "Male")) +
  scale_x_continuous( breaks = seq(1995, 2023, by = 5)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", colour = "Sex",
       title = "Annual incidence for JIA by Sex (1995-2023)") +
  theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"),legend.position = "right",
        legend.text = element_text(size = 11))

#View plot
JIA_plot

#Save plot to folder
#ggsave(JIA_plot, file = "R:/Anie/Incidence Plots/JIA/JIA_sex_incidence.png", width = 7, height = 5, dpi = 600)
```
## jSLE
```{r}
#Load relevant .xlsx file
jSLE_sex_graph <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual sex incidence rates.xlsx")
jSLE_sex_graph$Group <- factor(jSLE_sex_graph$Group, levels = c("F", "M")) 

#Plot jSLE annual sex incidence
#2023 was excluded as there were no cases but large CI - just like JDM
#Will have to make a table with 95% CI and add to supplementary/appendix
jSLE_plot <- ggplot(jSLE_sex_graph %>% filter(Year != 2023),
  aes(x = Year, y = IncidenceRate, colour = Group, group = Group)) +
  geom_line(linewidth = 1) + geom_point(size = 1) +
   scale_colour_viridis_d(labels = c("F" = "Female", "M" = "Male")) +
  scale_x_continuous(breaks = seq(1995, 2022, by = 5)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", colour = "Sex", title = "Annual incidence for jSLE by sex (1995-2022)") + theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"), legend.position = "right",
    legend.text = element_text(size = 11)
  )

#View plot
jSLE_plot

#Save plot to folder
#ggsave(jSLE_plot, file = "R:/Anie/Incidence Plots/jSLE/jSLE_sex_incidence.png", width = 7, height = 5, dpi = 600)
```

## Joining all sex graphs together
```{r}
#Using patchwork library for combined plot
combined_plot <- JIA_plot | jSLE_plot | JDM_plot

combined_plot

ggsave("combined_sex_plots.png",plot = combined_plot,width = 15,height = 5,units = "in",dpi = 300)
```

# Tables with incidence rates and 95% CI
## JDM
```{r}
#Setting wider format to reduce rows
JDM_wide <- JDM_sex_graph %>%
  mutate(Group = recode(Group, "F" = "Female", "M" = "Male"),
    `IR (95% CI)` = sprintf("%.2f (%.2f–%.2f)", IncidenceRate, Lower95CI, Upper95CI
    ), PersonYears = format(round(PersonYears), big.mark = ",")) %>%
  select(Year, Group, Cases, PersonYears,`IR (95% CI)`) %>%
  pivot_wider(names_from = Group, values_from = c(Cases, PersonYears, `IR (95% CI)`))

#Reorder columns
JDM_wide <- JDM_wide %>% select(Year, Cases_Female, PersonYears_Female, `IR (95% CI)_Female`, Cases_Male, PersonYears_Male, `IR (95% CI)_Male`)

#Rename columns
colnames(JDM_wide) <- c("Year", "Female Cases", "Female Person-years", "Female IR (95% CI)", "Male Cases", "Male Person-years", "Male IR (95% CI)")

#Write to csv for printing
write_xlsx(JDM_wide, "Annual sex-specific JDM w 95% CI.xlsx")

#Print using kable
knitr::kable(JDM_wide, caption = "Annual incidence of JDM by sex, 1995–2022.",
  align = c("c", "r", "r", "c", "r", "r", "c"), booktabs = TRUE)
```
##jSLE
```{r}
#Setting wider format to reduce rows
jSLE_wide <- jSLE_sex_graph  %>%
  mutate(Group = recode(Group, "F" = "Female", "M" = "Male"),
    `IR (95% CI)` = sprintf("%.2f (%.2f–%.2f)", IncidenceRate, Lower95CI, Upper95CI
    ), PersonYears = format(round(PersonYears), big.mark = ",")) %>%
  select(Year, Group, Cases, PersonYears,`IR (95% CI)`) %>%
  pivot_wider(names_from = Group, values_from = c(Cases, PersonYears, `IR (95% CI)`))

#Reorder columns
jSLE_wide <- jSLE_wide %>% select(Year, Cases_Female, PersonYears_Female, `IR (95% CI)_Female`, Cases_Male, PersonYears_Male, `IR (95% CI)_Male`)

#Rename columns
colnames(jSLE_wide) <- c("Year", "Female Cases", "Female Person-years", "Female IR (95% CI)", "Male Cases", "Male Person-years", "Male IR (95% CI)")

#Write to csv for printing
write_xlsx(jSLE_wide, "Annual sex-specific jSLE w 95% CI.xlsx")

#Print using kable
knitr::kable(jSLE_wide, caption = "Annual incidence of jSLE by sex, 1995–2022.",
  align = c("c", "r", "r", "c", "r", "r", "c"), booktabs = TRUE)
```
## JIA
```{r}
#Setting wider format to reduce rows
JIA_wide <- JIA_sex_graph %>%
  mutate(Group = recode(Group, "F" = "Female", "M" = "Male"),
    `IR (95% CI)` = sprintf("%.2f (%.2f–%.2f)", IncidenceRate, Lower95CI, Upper95CI
    ), PersonYears = format(round(PersonYears), big.mark = ",")) %>%
  select(Year, Group, Cases, PersonYears,`IR (95% CI)`) %>%
  pivot_wider(names_from = Group, values_from = c(Cases, PersonYears, `IR (95% CI)`))

#Reorder columns
JIA_wide <- JIA_wide %>% select(Year, Cases_Female, PersonYears_Female, `IR (95% CI)_Female`, Cases_Male, PersonYears_Male, `IR (95% CI)_Male`)

#Rename columns
colnames(JIA_wide) <- c("Year", "Female Cases", "Female Person-years", "Female IR (95% CI)", "Male Cases", "Male Person-years", "Male IR (95% CI)")

#Write to csv for printing
write_xlsx(JIA_wide, "Annual sex-specific JIA w 95% CI.xlsx")

#Print using kable
knitr::kable(JIA_wide, caption = "Annual sex-specific incidence of JIA, 1995–2022.",
  align = c("c", "r", "r", "c", "r", "r", "c"), booktabs = TRUE)
```

# Heat map of Annual sex incidence rates
## JDM
```{r}
#Plot heatmap
ggplot(JDM_sex_graph, aes(x=factor(Year),
                                   y=factor(Group),
                                   fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, na.value = "grey90", name = "Incidence\nper 100,000") + #na values appear grey - 2023, NI
  labs( x = "Year", y = "Sex", title = "Annual Sex Incidence Rates of JDM (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(face = "bold"))
```
## jSLE
```{r}
#Plot heatmap
ggplot(jSLE_sex_graph, aes(x=factor(Year),
                                   y=factor(Group),
                                   fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, na.value = "grey90", name = "Incidence\nper 100,000") + #na values appear grey - 2023, NI
  labs( x = "Year", y = "Sex", title = "Annual Sex Incidence Rates of jSLE (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(face = "bold"))

```
## JIA
```{r}
#Load in JIA data
JIA_sex_graph <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA incidence rate - overall,sex,region.xlsx", sheet = 2)

#Plot heatmap
ggplot(JIA_sex_graph, aes(x=factor(Year),  y=factor(Group), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, na.value = "grey90", name = "Incidence\nper 100,000") + #na values appear grey - 2023, NI
  labs( x = "Year", y = "Sex", title = "Annual Sex Incidence Rates of JIA (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(),
        plot.title = element_text(face = "bold"))
```

# Heat Map of Annual regional incidence rates
## JIA
```{r}
#Read in csv file
JIA_regional_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA Incidence rate - overall,sex,region.xlsx", sheet = 3)

#Plot heatmap
ggplot(JIA_regional_incidence, aes(x=factor(Year), y=factor(Group), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, name = "Incidence\nper 100,000") + 
  labs( x = "Year", y = "Region of the UK", title = "Annual Regional Incidence Rates of JIA (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(), plot.title = element_text(face = "bold"))

#Save heatmap
ggsave("R:/Anie/Incidence Plots/JIA/JIA_regional_heatmap.png", width = 14, height = 8, dpi = 600)
```
## JDM
```{r}
#Read in csv file
JDM_regional_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual region incidence rates.xlsx")

#Plot heatmap
ggplot(JDM_regional_incidence, aes(x=factor(Year), y=factor(Group), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, name = "Incidence\nper 100,000") +
  labs( x = "Year", y = "Region of the UK", title = "Annual Regional Incidence Rates of JDM (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank(), plot.title = element_text(face = "bold"))

#Save plot
ggsave("R:/Anie/Incidence Plots/JDM/JDM_heatmap.png", width = 14, height = 8, dpi = 600)
```
## jSLE
```{r}
#Read in csv file
jSLE_regional_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual region incidence rates.xlsx")

#Plot heatmap
ggplot(jSLE_regional_incidence, aes(x=factor(Year), y=factor(Group), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, name = "Incidence\nper 100,000") + 
  labs( x = "Year", y = "Region of the UK", title = "Annual Regional Incidence Rates of jSLE (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid = element_blank(),
        plot.title = element_text(face = "bold"))

#Save plot
ggsave("R:/Anie/Incidence Plots/jSLE/jSLE_heatmap.png", width = 14, height = 8, dpi = 600)
```

# Line graphs for annual regional incidence rates
## JIA
```{r}
#Load relevant .xlsx file
JIA_regional_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA Incidence rate - overall,sex,region.xlsx", sheet = 3)

#Plot JIA annual sex incidence
#Will have to make a table with 95% CI and add to supplementary/appendix
JIA_plot <- ggplot(JIA_regional_incidence,
                   aes(x = Year, y = IncidenceRate, colour = factor(Group), group = factor(Group))) +
  geom_line(linewidth = 0.9) + geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(1995, 2022, by = 5)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", colour = "Region", 
    title = "Annual Regional Incidence for JIA (1995-2023)") + theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"), legend.position = "right",
        legend.text = element_text(size = 11))

#View plot
JIA_plot

#Save plot to folder
#ggsave("R:/Anie/Incidence Plots/JIA/JIA_regional_incidence_line.png", width = 7, height = 5, dpi = 600)
```
## JDM
```{r}
#Load relevant .xlsx file
JDM_regional_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual region incidence rates.xlsx")

#Plot JDM annual sex incidence
#Will have to make a table with 95% CI and add to supplementary/appendix
JDM_plot <- ggplot(JDM_regional_incidence,
                   aes(x = Year, y = IncidenceRate, colour = factor(Group), group = factor(Group))) +
  geom_line(linewidth = 0.9) + geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(1995, 2022, by = 5)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years",
    colour = "Region", title = "Annual Regional Incidence for JDM (1995-2023)") +
  theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"),
    legend.position = "right", legend.text = element_text(size = 11))

#View plot
JDM_plot

#Save plot to folder
#ggsave("R:/Anie/Incidence Plots/JDM_regional_incidence_line.png", width = 7, height = 5, dpi = 600)
```
## jSLE
```{r}
#Load relevant .xlsx file
jSLE_regional_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual region incidence rates.xlsx")

#Plot JDM annual sex incidence
#Will have to make a table with 95% CI and add to supplementary/appendix
jSLE_plot <- ggplot(jSLE_regional_incidence,
                   aes(x = Year, y = IncidenceRate, colour = factor(Group), group = factor(Group))) +
  geom_line(linewidth = 0.9) + geom_point(size = 0.5) +
  scale_x_continuous(breaks = seq(1995, 2022, by = 5)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", colour = "Region",
    title = "Annual Regional Incidence for jSLE (1995-2023)") + theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"), legend.position = "right", 
        legend.text = element_text(size = 11))

#View plot
jSLE_plot

#Save plot to folder
#ggsave("R:/Anie/Incidence Plots/jSLE/jSLE_regional_incidence_line.png", width = 7, height = 5, dpi = 600)
```

# Building forest plots for overall regional incidence
## JIA
```{r}
JIA_overall_region <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/JIA Overall, sex and region results.xlsx")

#Filter overall and sex from xlsx file
JIA_overall_region <- JIA_overall_region %>% filter(Group == "region")

#Plot forest graph
ggplot(JIA_overall_region, aes(x = IncidenceRate, y = reorder(Level, IncidenceRate))) +
  geom_errorbarh(aes(xmin = Lower95CI, xmax = Upper95CI), height = 0.2, linewidth = 0.5, colour = "black") +
  geom_point(size = 2, colour = "red") +
  labs(title = "Overall Regional Incidence of JIA (1995-2023)", x = "Incidence per 100,000 person-years (95% CI)", y = "Region of the UK")+
  theme_minimal(base_size = 13) +
theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5))

#Save plot to folder Incidence Plots
#ggsave("JIA_region_forest.png",width = 8, height = 6, dpi = 300, bg = "white")
```
## JDM
```{r}
JDM_overall_region <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM Overall, by sex and by region.xlsx")

#Filter overall and sex from xlsx file
JDM_overall_region <- JDM_overall_region %>% filter(Group == "region")

#Plot forest graph
JDM_region_forest <- ggplot(JDM_overall_region,
       aes(x = IncidenceRate,y = reorder(Level, IncidenceRate))) +
  geom_errorbarh(aes(xmin = Lower95CI, xmax = Upper95CI), height = 0.2, linewidth = 0.5, colour = "black") +
  geom_point(size = 2, colour = "red") +
  labs(title = "Overall Regional Incidence of JDM (1995-2023)", x = "Incidence per 100,000 person-years (95% CI)", y = "Region of the UK")+
  theme_minimal(base_size = 13) +
theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5))

#Save plot to folder Incidence Plots
#ggsave("JDM_region_forest.png",width = 8, height = 6, dpi = 300, bg = "white")
```
## jSLE
```{r}
jSLE_overall_region <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE Overall, by sex and by region.xlsx")

#Exclude NA in following rows:
jSLE_overall_region <- jSLE_overall_region %>% filter(Group == "region")

#Plot forest graph
jSLE_region_forest <- ggplot(jSLE_overall_region, aes(x = IncidenceRate, y = reorder(Level, IncidenceRate))) +
  geom_errorbarh(aes(xmin = Lower95CI, xmax = Upper95CI), height = 0.2,
                 linewidth = 0.5, colour = "black") +
  geom_point(size = 2, colour = "red") +
  labs(title = "Overall Regional Incidence of jSLE (1995-2023)", 
       x = "Incidence per 100,000 person-years (95% CI)", y = "Region of the UK")+
  theme_minimal(base_size = 13) +
theme(panel.grid.major.y = element_blank(), panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5))

#Save plot to folder Incidence Plots
ggsave("jSLE_region_forest.png",width = 8, height = 6, dpi = 300, bg = "white")
```

```{r}
# Join the overall forest graphs together
combined_forest_region <- (JDM_region_forest | jSLE_region_forest) +
  plot_annotation(
    title = "Overall regional incidence of juvenile rheumatic diseases (1995–2023)",
    tag_levels = "A"
  )
combined_forest_region
ggsave(combined_forest_region, file = "JDM and jSLE Overall regional incidence forest.png", width = 14,height = 4,dpi = 300)

JDM_plot <- JDM_overall_region %>%
  mutate(Disease = "JDM")

jSLE_plot <- jSLE_overall_region %>%
  mutate(Disease = "jSLE")

combined <- bind_rows(JDM_plot, jSLE_plot)

ggplot(combined,aes(x = IncidenceRate,y = Level,xmin = Lower95CI,xmax = Upper95CI)) +
  geom_errorbarh(height = 0.15) +
  geom_point(size = 2.5, colour = "red") +
  facet_wrap(~Disease, scales = "free_x") +
  labs(title = "Overall crude regional incidence rate of juvenile rheumatic diseases (1995–2023)", x = "Incidence per 100,000 person-years (95% CI)", y = "Region of the UK") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), strip.text = element_text(face = "bold", size = 12), axis.text.y = element_text(size = 10), panel.spacing = unit(1.2, "lines"))

ggsave("jSLE and JDM overall regional incidence_forest.png", width = 8, height = 6, dpi = 300, bg = "white")
```
# Heat Map of Annual age-specific incidence rates
## JIA
```{r}
#Read in csv file
JIA_annual_age <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/JIA annual age-specific incidence.xlsx")

#Plot heatmap
ggplot(JIA_annual_age[JIA_annual_age$Age != 0, ], aes(x=factor(Age), y=factor(Year), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, name = "Incidence\nper 100,000") +
  labs( x = "Age at Diagnosis (years)", y = "Year",
        title = "Annual Age-specific Incidence Rates of JIA (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(hjust = 1),
        panel.grid = element_blank(), plot.title = element_text(face = "bold"))

#Save heatmap
#ggsave("JIA_annual_age_heatmap.png", width = 14, height = 8, dpi = 600)
```
## JDM
```{r}
#Read in csv file
JDM_annual_age <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual age-specific incidence rates.xlsx")
#Plot heatmap
ggplot(JDM_annual_age[JDM_annual_age$Age != 0, ], aes(x=factor(Age), y=factor(Year), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, name = "Incidence\nper 100,000") + 
  labs( x = "Age at Diagnosis (years)", y = "Year",
        title = "Annual Age-specific Incidence Rates of JDM (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(hjust = 1),
        panel.grid = element_blank(), plot.title = element_text(face = "bold"))

#Save heatmap
ggsave("JDM_annual_age_heatmap.png", width = 14, height = 8, dpi = 600)
```
## jSLE
```{r}
#Read in csv file
jSLE_annual_age <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual_age-specific.xlsx")

#Plot heatmap
ggplot(jSLE_annual_age[jSLE_annual_age$Age != 0, ], aes(x=factor(Age), y=factor(Year), fill = IncidenceRate)) +
  geom_tile(colour = "white", linewidth = 0.1) +
  scale_fill_viridis(option = "viridis", direction = 1, name = "Incidence\nper 100,000") +
  labs( x = "Age at Diagnosis (years)", y = "Year",
        title = "Annual Age-specific Incidence Rates of jSLE (1995-2023)") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(hjust = 1), panel.grid = element_blank(),
        plot.title = element_text(face = "bold"))

#Save heatmap
ggsave("jSLE_annual_age_heatmap.png", width = 14, height = 8, dpi = 600)
```

# Choropleth maps for regional overall
## JDM
```{r}
#Read SHA geographical boundaries
england_regions <- st_read("R:/Anie/Incidence Plots/Regions_December_2023_Boundaries_EN_BUC_-8646452588026772277/RGN_DEC_2023_EN_BUC.shp")

#Read in region_incidence for JDM
JDM_overall_region <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM Overall, by sex and by region.xlsx")

#Removing sex, overall from xlsx file
region_incidence <- JDM_overall_region %>% filter(Group == "region")

#Remove NI - didn't make much difference
region_incidence <- region_incidence %>% filter(Level != "Northern Ireland")

#Rename column name to match england_regions
region_incidence <- region_incidence %>% rename(Region = Level)

#Do left join to join shapefile and data
map_data <- england_regions %>% left_join(region_incidence, by = c("RGN23NM" = "Region"))

#Plot choropleth
JDM_choropleth <- ggplot(map_data) +
  geom_sf(aes(fill = IncidenceRate), colour = "white", size = 0.1) +
  scale_fill_viridis_c(option = "viridis", name = "Incidence\nper 100,000 person-years") +
  labs(title = "Regional Incidence of JDM (1995 - 2023)", caption = "Data: CPRD Aurum") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        panel.grid = element_blank(), axis.text = element_blank(), 
        axis.ticks = element_blank(), axis.title = element_blank())

#print map
JDM_choropleth

#Save choropleth map
#ggsave(JDM_choropleth, file ="JDM_overall_choropleth.png")
```
## jSLE
```{r}
#Read SHA geographical boundaries
england_regions <- st_read("R:/Anie/Incidence Plots/Regions_December_2023_Boundaries_EN_BUC_-8646452588026772277/RGN_DEC_2023_EN_BUC.shp")

#Read in region_incidence for jSLE
jSLE_overall_region <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE Overall, by sex and by region.xlsx")

#Removing sex, overall from xlsx file
region_incidence <- jSLE_overall_region %>% filter(Group == "region")

#Remove NI - didn't make much difference
region_incidence <- region_incidence %>% filter(Level != "Northern Ireland")

#Rename column name to match england_regions
region_incidence <- region_incidence %>% rename(Region = Level)

#Do left join to join shapefile and data
map_data <- england_regions %>% left_join(region_incidence, by = c("RGN23NM" = "Region"))

#Plot choropleth
jSLE_choropleth <- ggplot(map_data) +
  geom_sf(aes(fill = IncidenceRate), colour = "white", size = 0.1) +
  scale_fill_viridis_c(option = "viridis", name = "Incidence\nper 100,000 person-years") +
  labs(title = "Regional Incidence of jSLE (1995 - 2023)",caption = "Data: CPRD Aurum") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5),
        panel.grid = element_blank(), axis.text = element_blank(), 
        axis.ticks = element_blank(), axis.title = element_blank())

#print map
jSLE_choropleth

#Save choropleth map
#ggsave(jSLE_choropleth, file = "jSLE_overall_choropleth.png")
```
## JIA
```{r}
#Read SHA geographical boundaries
england_regions <- st_read("R:/Anie/Incidence Plots/Regions_December_2023_Boundaries_EN_BUC_-8646452588026772277/RGN_DEC_2023_EN_BUC.shp")

#Read in region_incidence for JDM
JIA_overall_region <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/JIA Overall, sex and region results.xlsx")

#Removing sex, overall from xlsx file
region_incidence <- JIA_overall_region %>% filter(Group == "region")

#Remove NI - didn't make much difference
region_incidence <- region_incidence %>% filter(Level != "Northern Ireland")

#Rename column name to match england_regions
region_incidence <- region_incidence %>% rename(Region = Level)

#Do left join to join shapefile and data
map_data <- england_regions %>% left_join(region_incidence, by = c("RGN23NM" = "Region"))

#Plot choropleth
JIA_choropleth <- ggplot(map_data) +
  geom_sf(aes(fill = IncidenceRate), colour = "white", size = 0.1) +
  scale_fill_viridis_c(option = "viridis", name = "Incidence\nper 100,000 person-years") +
  labs(title = "Regional Incidence of JIA (1995 - 2023)",caption = "Data: CPRD Aurum") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), 
        panel.grid = element_blank(), axis.text = element_blank(), 
        axis.ticks = element_blank(),axis.title = element_blank())

#print map
JIA_choropleth

#Save choropleth map
ggsave(JIA_choropleth, file = "JIA_overall_choropleth.png")
```

# Bar chart of overall incidence by age
## JDM
```{r}
#Load up JDM age-specific overall 
JDM_overall_age <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM overall age-specific.xlsx")

#Bar Chart
JDM_overall_age_bar <- ggplot(JDM_overall_age, aes(x = factor(Age), y = IncidenceRate)) +
  geom_col(fill = "pink", colour = "black", linewidth = 0.3) +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.25, linewidth = 0.5) +
  labs(x = "Age at diagnosis (years)", y = "Incidence rate per 100,000 person-years",
    title = "Overall Age-specific incidence of JDM (1995–2023)") + theme_classic(base_size = 12)

#Point plot
ggplot(JDM_overall_age[JDM_overall_age$Age != 0, ], aes(x = Age, y = IncidenceRate)) +
  geom_point(size = 2, colour = "red") + geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI),
    width = 0.15) +
  scale_x_continuous(breaks = 0:18) +
  labs(title = "Overall JDM incidence by Age at Diagnosis",x = "Age at diagnosis (years)",y = "Incidence rate per 100,000 person-years") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggsave("JDM_overall_age_point plot.png", width = 8, bg = "white")
```
## JIA 
```{r}
#Load in JIA overall age incidence
JIA_overall_age <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/JIA overall age incidence.xlsx")

names(JIA_overall_age)

JIA_overall_age <- JIA_overall_age[Age!= 0]

#Plot bar chart
JIA_overall_age_bar <- ggplot(JIA_overall_age[JIA_overall_age$Age != 0, ], aes(x = factor(Age), y = IncidenceRate)) +
  geom_col(fill = "lightblue", colour = "black", linewidth = 0.3) +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.25, linewidth = 0.5) +
  labs(x = "Age at diagnosis (years)", y = "Incidence rate per 100,000 person-years",
    title = "Overall Age-specific incidence of JIA (1995–2023)") +theme_classic(base_size = 12)


ggplot(JIA_overall_age[JIA_overall_age$Age != 0, ], aes(x = Age, y = IncidenceRate)) +
  geom_point(size = 2, colour = "red") + geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI),
    width = 0.15) +
  scale_x_continuous(breaks = 0:16) +
  labs(title = "Overall JIA incidence by Age at Diagnosis",x = "Age at diagnosis (years)",y = "Incidence rate per 100,000 person-years") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

#Save plot
ggsave("JIA overall age point plot.png", bg = "white", width = 8)
```
## jSLE
```{r}
#Load up jSLE overall age-specific incidence
jSLE_overall_age <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE Overall by age.xlsx")

#Plot barchart
jSLE_overall_age_bar <- ggplot(jSLE_overall_age, aes(x = factor(Age), y = IncidenceRate)) +
  geom_col(fill = "salmon", colour = "black", linewidth = 0.3) +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.25, linewidth = 0.5) +
  labs(x = "Age at diagnosis (years)", y = "Incidence rate per 100,000 person-years",
    title = "Overall Age-specific incidence of jSLE (1995–2023)") +theme_classic(base_size = 11)

#Point plot
ggplot(jSLE_overall_age[jSLE_overall_age$Age != 0, ], aes(x = Age, y = IncidenceRate)) +
  geom_point(size = 2, colour = "red") + geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI),
    width = 0.15) +
  scale_x_continuous(breaks = 0:18) +
  labs(title = "Overall jSLE incidence by Age at Diagnosis",x = "Age at diagnosis (years)",y = "Incidence rate per 100,000 person-years") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggsave("jSLE_overall_age_point plot.png", width = 8, bg = "white")

#Save plot
#ggsave(jSLE_overall_age_bar, file = "jSLE overall age bar.png")
```

# Line graph of age-specific incidence rates
## JDM
```{r}
ggplot(JDM_overall_age, aes(x = Age, y = IncidenceRate)) +
  geom_line(linewidth = 0.8, colour = "red") +
  geom_point(size = 2, colour = "red") +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.2, linewidth = 0.3) +
  scale_x_continuous(breaks = 0:18) +
  labs(title = "Age-specific incidence of JDM (1995–2023)", x = "Age at diagnosis (years)",
    y = "Incidence rate per 100,000 person-years") +theme_classic(base_size = 11)
```
## JIA
```{r}
ggplot(JIA_overall_age, aes(x = Age, y = IncidenceRate)) +
  geom_line(linewidth = 0.8, colour = "red") +
  geom_point(size = 2, colour = "red") +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.2, linewidth = 0.3) +
  scale_x_continuous(breaks = 0:18) +
  labs(title = "Age-specific incidence of JIA (1995–2023)",x = "Age at diagnosis (years)",
    y = "Incidence rate per 100,000 person-years") + theme_classic(base_size = 11)
```
## jSLE
```{r}
#Plot line graph
ggplot(jSLE_overall_age, aes(x = Age, y = IncidenceRate)) +
  geom_line(linewidth = 0.8, colour = "red") +
  geom_point(size = 2, colour = "red") +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.2, linewidth = 0.3) +
  scale_x_continuous(breaks = 0:18) +
  labs(title = "Age-specific incidence of jSLE (1995–2023)", x = "Age at diagnosis (years)",
    y = "Incidence rate per 100,000 person-years") + theme_classic(base_size = 11)
```

# Annual incidence rate bar chart
## JIA
```{r}
#Load up JIA annual incidence 
JIA_annual_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA Incidence rate - overall,sex,region.xlsx", sheet = 1)

JIA_plot <- ggplot(JIA_annual_incidence,
                   aes(x = Year, y = IncidenceRate))+
  geom_line(linewidth = 0.9, colour = "darkgreen") + geom_point(size = 1, colour = "darkgreen") +
  scale_x_continuous(breaks = seq(1995, 2023, by = 2)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", 
    title = "Annual Incidence for JIA (1995-2023)") + theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"), legend.position = "right",
        legend.text = element_text(size = 11),
        theme(axis.text.x = element_text(angle = 45, hjust = 1)))

#View plot
JIA_plot

ggsave("JIA_annual_incidence.png", bg = "white", width = 8)
```
## JDM
```{r}
#Load up JDM annual incidence
JDM_annual_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual incidence rates.xlsx")

JDM_plot <- ggplot(JDM_annual_incidence,
                   aes(x = Year, y = IncidenceRate))+
  geom_line(linewidth = 0.9, colour = "darkblue") + geom_point(size = 1, colour = "darkblue") +
  scale_x_continuous(breaks = seq(1995, 2023, by = 2)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", 
    title = "Annual Incidence for JDM (1995-2023)") + theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"), legend.position = "right",
        legend.text = element_text(size = 11),
        theme(axis.text.x = element_text(angle = 45, hjust = 1)
        ))

#View plot
JDM_plot

ggsave("JDM_annual_incidence.png", bg = "white", width = 8)
```
## jSLE
```{r}
#Load up JDM annual incidence
jSLE_annual_incidence <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual incidence rates.xlsx")

jSLE_plot <- ggplot(jSLE_annual_incidence,
                   aes(x = Year, y = IncidenceRate))+
  geom_line(linewidth = 0.9, colour = "darkorange") + geom_point(size = 1, colour = "darkorange") +
  scale_x_continuous(breaks = seq(1995, 2023, by = 2)) +
  labs(x = "Calendar year", y = "Incidence rate per 100,000 person-years", 
    title = "Annual Incidence for jSLE (1995-2023)") + theme_classic(base_size = 12) +
  theme(axis.text = element_text(colour = "black"), legend.position = "right",
        legend.text = element_text(size = 11), )

#View plot
jSLE_plot


# Join the annual sex line graphs together
library(patchwork)

combined_plot_annual <- JDM_plot | jSLE_plot
combined_plot_annual

ggsave(combined_plot_annual, file = "Annual incidence JDM and jSLE.png", width = 14,height = 4,dpi = 300)
```

# Line graph (w/ 95% CI) for point prevalence
## JIA
```{r}
JIA_pp <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Point Prevalence for JIA.xlsx")

ggplot(JIA_pp, aes(x = Year, y = Prevalence)) +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.4, colour = "black") +
  geom_line(linewidth = 0.8, color = "red") +
  geom_point(size = 1.5, colour = "red") +
  scale_x_continuous(breaks = 1995:2022)+
  labs(title = "Annual point prevalence of JIA",x = "Year",y = "Prevalence per 100,000") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_minimal()

ggsave("Annual point prevalence of JIA.png", bg = "white", width = 14, height=7, dpi = 300)
```
## JDM
```{r}
JDM_pp <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/Point Prevalence for JDM.xlsx")

ggplot(JDM_pp, aes(x = Year, y = Prevalence)) +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.4, colour = "black") +
  geom_line(linewidth = 0.8, color = "red") +
  geom_point(size = 1.5, colour = "red") +
  scale_x_continuous(breaks = 1995:2022)+
  labs(title = "Annual point prevalence of JDM",x = "Year",y = "Prevalence per 100,000") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_minimal()

ggsave("Annual point prevalence of JDM.png", bg = "white", width = 14, height=7, dpi = 300)
```
## jSLE
```{r}
jSLE_pp <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/Point Prevalence for jSLE.xlsx")

ggplot(jSLE_pp, aes(x = Year, y = Prevalence)) +
  geom_errorbar(aes(ymin = Lower95CI, ymax = Upper95CI), width = 0.4, colour = "black") +
  geom_line(linewidth = 0.8, color = "red") +
  geom_point(size = 1.5, colour = "red") +
  scale_x_continuous(breaks = 1995:2022)+
  labs(title = "Annual point prevalence of jSLE",x = "Year",y = "Prevalence per 100,000") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme_minimal()

ggsave("Annual point prevalence of jSLE.png", bg = "white", width = 14, height=7, dpi = 300)
```