---
title: "Poisson Regression/Negative binomial regression"
author: "Anie Thomas"
date: "2026-08-02"
output: html_document
---

Poisson regression models the number of cases while accounting for the amount of person-time through an offset.
All graphs were built using the results obtained from Incidence_Analysis.R

# Libraries
```{r}
library(readxl)
#install.packages("MASS")   # only once
library(MASS)
library(knitr)
library(kableExtra)
library(writexl)
```

# 1. JIA

## JIA Sex differences
```{r}
#Read in JIA annual sex
JIA_annual_sex <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA Incidence rate - overall,sex,region.xlsx", sheet = 2)

JIA_annual_sex$Group <- factor(JIA_annual_sex$Group)

JIA_model_sex <- glm(Cases ~ Group, offset = log(PersonYears), family = poisson(link="log"),
                     data = JIA_annual_sex)

summary(JIA_model_sex)

#Obtain IRR
exp(coef(JIA_model_sex))

#95% CI
exp(confint(JIA_model_sex))

#AIC = 1719.5, Residual deviance: 1311.1 and 56 degrees of freedom

#Assess overdispersion
pearson_dispersion <- sum(residuals(JIA_model_sex, type = "pearson")^2) /
  df.residual(JIA_model_sex)

pearson_dispersion #22 indicates overdispersion
```
## Negative Binomial Regression for sex
```{r}
#Using negative binomial regression - overdispersion seen
JIA_model_sex_nb <- glm.nb(Cases ~ Group + Year + offset(log(PersonYears)),
  data = JIA_annual_sex)

summary(JIA_model_sex_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(JIA_model_sex_nb), confint(JIA_model_sex_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(JIA_model_sex_nb), confint(JIA_model_sex_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)

#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of JIA sex from the negative binomial regression model."
)

```

## Regional differences
```{r}
#Read in annual regional incidence
JIA_annual_region <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/Annual JIA Incidence rate - overall,sex,region.xlsx", sheet=3)

#Factorise by group
JIA_annual_region$Group <- factor(JIA_annual_region$Group)

#Removal of NI 2023 (0 cases, 0 person-years)
JIA_annual_region_clean <- subset(JIA_annual_region,!(Group == "Northern Ireland" & Year == 2023))

#Poisson Regression
JIA_model_region <- glm(Cases ~ Group, offset = log(PersonYears), family = poisson(link="log"),
                     data = JIA_annual_region_clean)

#Assess residual variance and degrees of freedom - evidence of over dispersion (1818.8 on 279 df)
summary(JIA_model_region)

#Obtain IRR
exp(coef(JIA_model_region))

#95% CI
exp(confint(JIA_model_region))
```
## Negative binomial model for region
```{r}
JIA_model_region_nb <- glm.nb(Cases ~ Group + Year + offset(log(PersonYears)), data = JIA_annual_region_clean)

#East Midlands is the reference
summary(JIA_model_region_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(JIA_model_region_nb), confint(JIA_model_region_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(JIA_model_region_nb), confint(JIA_model_region_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)

#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of JIA region from the negative binomial regression model.")
```
## Age differences
```{r}
#Read in annual age incidence
JIA_annual_age <- read_excel("R:/Anie/Incidence xlsx sheets/JIA/JIA annual age-specific incidence.xlsx")

#Fit model
JIA_model_age_nb <- glm.nb(Cases ~ factor(Age) + Year + offset(log(PersonYears)), data = JIA_annual_age)

#Summary - age(0) is the reference
summary(JIA_model_age_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(JIA_model_age_nb), confint(JIA_model_age_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(JIA_model_age_nb), confint(JIA_model_age_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)


#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of JIA age at diagnosis from the negative binomial regression model.")
```

# 2. JDM
## Sex differences
```{r}
#Read in JDM annual sex
JDM_annual_sex <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual sex incidence rates.xlsx")

# JIA_annual_sex$Group <- factor(JIA_annual_sex$Group)

JDM_model_sex <- glm(Cases ~ factor(Group), offset = log(PersonYears), family = poisson(link="log"),
                     data = JDM_annual_sex)

summary(JDM_model_sex)

#Obtain IRR
exp(coef(JDM_model_sex))

#95% CI
exp(confint(JDM_model_sex))

#AIC = 278.99, Residual deviance: 79.907 and 56 degrees of freedom

#Assess overdispersion
pearson_dispersion <- sum(residuals(JDM_model_sex, type = "pearson")^2) /
  df.residual(JDM_model_sex)

pearson_dispersion #1.48 indicates overdispersion - will do negative binomial models

#Ratio > 1 is overdispersion
#Ratio < 1 is underdispersion
#Ratio ~ 1 is dispersion consistent
```

```{r}
#Using negative binomial regression - overdispersion seen
JDM_model_sex_nb <- glm.nb(Cases ~ factor(Group) + Year + offset(log(PersonYears)),
  data = JDM_annual_sex)

summary(JDM_model_sex_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(JDM_model_sex_nb), confint(JDM_model_sex_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(JDM_model_sex_nb), confint(JDM_model_sex_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)

#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of JDM sex from the negative binomial regression model."
)

```

## Regional differences
```{r}
#Load up data
JDM_annual_region <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual region incidence rates.xlsx")

#Removal of NI 2023 (0 cases, 0 person-years)
JDM_annual_region_clean <- subset(JDM_annual_region,!(Group == "Northern Ireland" & Year == 2023))

#Negative binomial model
JDM_model_region_nb <- glm.nb(Cases ~ factor(Group) + Year + offset(log(PersonYears)), data = JDM_annual_region_clean)

#East Midlands is the reference
summary(JDM_model_region_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(JDM_model_region_nb), confint.default(JDM_model_region_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(JDM_model_region_nb), confint.default(JDM_model_region_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)

#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of JDM region from the negative binomial regression model.")
```

## Age differences
```{r}
#Read in annual age incidence
JDM_annual_age <- read_excel("R:/Anie/Incidence xlsx sheets/JDM/JDM annual age-specific incidence rates.xlsx")

#Fit model
JDM_model_age_nb <- glm.nb(Cases ~ factor(Age) + Year + offset(log(PersonYears)), data = JDM_annual_age)

#Summary - age(0) is the reference
summary(JDM_model_age_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(JDM_model_age_nb), confint(JDM_model_age_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(JDM_model_age_nb), confint(JDM_model_age_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)


#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of JDM age at diagnosis from the negative binomial regression model.")
```


# 3. jSLE
## Sex differences
```{r}
#Read in JDM annual sex
jSLE_annual_sex <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual sex incidence rates.xlsx")

jSLE_model_sex <- glm(Cases ~ factor(Group), offset = log(PersonYears), family = poisson(link="log"),
                     data = jSLE_annual_sex)

summary(jSLE_model_sex)

#Obtain IRR
exp(coef(jSLE_model_sex))

#95% CI
exp(confint(jSLE_model_sex))

#AIC = 321.36, Residual deviance: 88.78 and 56 degrees of freedom

#Assess overdispersion
pearson_dispersion <- sum(residuals(jSLE_model_sex, type = "pearson")^2) /
  df.residual(jSLE_model_sex)

pearson_dispersion #1.49 indicates overdispersion - will do negative binomial models

#Ratio > 1 is overdispersion
#Ratio < 1 is underdispersion
#Ratio ~ 1 is dispersion consistent
```

```{r}
#Using negative binomial regression - overdispersion seen
jSLE_model_sex_nb <- glm.nb(Cases ~ factor(Group) + Year + offset(log(PersonYears)),
  data = jSLE_annual_sex)

summary(jSLE_model_sex_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(jSLE_model_sex_nb), confint(jSLE_model_sex_nb)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(jSLE_model_sex_nb), confint(jSLE_model_sex_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)

#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of jSLE sex from the negative binomial regression model."
)

```

## Regional differences
```{r}
#Load up data
jSLE_annual_region <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual region incidence rates.xlsx")

#Removal of NI 2023 (0 cases, 0 person-years)
jSLE_annual_region_clean <- subset(jSLE_annual_region,!(Group == "Northern Ireland" & Year == 2023))

#Negative binomial model
jSLE_model_region_nb <- glm.nb(Cases ~ factor(Group) + Year + offset(log(PersonYears)), data = jSLE_annual_region_clean)

#East Midlands is the reference
summary(jSLE_model_region_nb)

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(jSLE_model_region_nb), confint(jSLE_model_region_nb))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)

#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of jSLE region from the negative binomial regression model.")
```
## Age differences
```{r}
#Read in annual age incidence
jSLE_annual_age <- read_excel("R:/Anie/Incidence xlsx sheets/jSLE/jSLE annual_age-specific.xlsx")

#Fit Poisson Model for jSLE age
jSLE_model_age <- glm(Cases ~ factor(Age), offset = log(PersonYears), family = poisson(link="log"),
                     data = jSLE_annual_age)

#Summary of Poisson model
summary(jSLE_model_age)

#Obtain IRR
exp(coef(jSLE_model_age))

#95% CI
exp(confint(jSLE_model_age))

#AIC = 1328.8, Residual deviance: 88.78 and 56 degrees of freedom

#Assess overdispersion
pearson_dispersion <- sum(residuals(jSLE_model_age, type = "pearson")^2) /
  df.residual(jSLE_model_age)

pearson_dispersion #1.03 indicates consistent dispersion - will do poisson

#Ratio > 1 is overdispersion
#Ratio < 1 is underdispersion
#Ratio ~ 1 is dispersion consistent

# #Fit model
# jSLE_model_age_nb <- glm.nb(Cases ~ factor(Age) + Year + offset(log(PersonYears)), data = jSLE_annual_age)
# 
# #Summary - age(0) is the reference
# summary(JDM_model_age_nb)

#Extract IRR and 95% CI
exp(cbind(IRR = coef(jSLE_model_age), confint(jSLE_model_age)))

#Put results into a table
#Extract IRRs and CIs
IRR_table <- as.data.frame(exp(cbind(
    IRR = coef(jSLE_model_age), confint(jSLE_model_age))))

#Add variable names
IRR_table$Variable <- rownames(IRR_table)


#Clean up names of row
IRR_table$Variable <- gsub("^Group", "", IRR_table$Variable)
IRR_table$Variable <- gsub("^Year$", "Calendar year", IRR_table$Variable)

#Create IRR(95% CI) column
IRR_table$`Adjusted IRR (95% CI)` <- sprintf("%.2f (%.2f–%.2f)", IRR_table$IRR,
          IRR_table$`2.5 %`, IRR_table$`97.5 %`)

# Keep only required columns
IRR_table <- IRR_table[, c("Variable", "Adjusted IRR (95% CI)")]

#Print table
kable(IRR_table, booktabs = TRUE,
  caption = "Adjusted incidence rate ratios of jSLE age at diagnosis from the negative binomial regression model.")
```